Einstein's gravity from a polynomial affine model

### 2016-05-04 AffGR-PRD.tex
Almost complete. Please finish conclusions

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds ---with or without cosmological constant--- as solutions. Moreover, the effective field equations coincide with some of those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity with a ``minimally'' coupled scalar field in this affine model. Finally, we present a naive analysis of the propagators of the gravitational theory, an their relation with the presence of ghost in the model. 
\end{abstract}

@@

However, General Relativity seems to be an effective theory of a, yet to be discovered, more fundamental one. A first hint was pointed out by E.~Cartan, after remarking that the curvature is not built up from a metric field but from a connection, and that a part of it was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's approach in General Relativity, in which both metric and connection fields are independent, and noticing that generically the theory possesses terms which do not appear in General Relativity. In addition, we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization, General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions ---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, by Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

%% Classically, General Relativity is the most successful theory of gravitational interactions. However, it is believed to be the effective theory of a, yet to be discovered, more fundamental one. Among the problems of General Relativity we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions ---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, and Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Equation~\eqref{SimpleEOM} is a generalization of a condition known in Riemannian geometry as covariantly constant Ricci curvature ---aka parallel Ricci curvature---, \mbox{$\nab{\rho} R_{\mu\nu} = 0$.}  All Einstein manifolds, whose Ricci tensor is proportional to the metric, \mbox{$R_{\mu\nu} \propto g_{\mu\nu}$,} satisfy the parallel Ricci condition due to the metricity condition, and therefore every vacuum solution to the Einstein's equations solves the (torsion-free) field equations of our model. Consequently, the fact that the non-relativistic limit of the gravitational potential in Eq.~\eqref{NewtonPot} yields a Newtonian potential seems clearer, and we can argue that even the post-Newtonian corrections coming from General Relativity are present in the chosen scenario of our model. An additional comment, the parallel Ricci curvature assures that although the manifold is not Einstein in general, the metric has to be locally a product of Einstein metrics~\cite{Besse}.

@@

Interestingly, the Eq.~\eqref{SimpleEOM} ---equivalently Eq.~\eqref{harm-curv}--- can be obtained from an effective gravitational Yang--Mills functional action~\cite{stephenson1958quadratic,kilmister1961use,Yang1974}, %constructed as the trace of the 
\begin{equation}
  \label{SKY}
  S_{\textsc{ym}} = \int \Tr \left( \Rif{}{} \st \Rif{}{} \right) = \int \left( \Rif{a}{b} \st \Rif{b}{a} \right),
\end{equation}
where $\Rif{}{} \in \Omega^2(\Mi, T^*\Mi \otimes T\Mi)$ is the curvature two-form, the operator $\st$ denotes the Hodge star, and the trace is taken on the bundle indices (see Ref.~\cite{bourguignon1982yang}).

@@

Until now, the most general diffeomorphism-invariant and power-counting renormalizable (gravitational) theory for an affine connection has been built, and we have showed that in a certain sector it is equivalent to General Relativity. In what follows, we show an attempt of including scalar matter into the model. The theory does not require a spacetime metric, but instead we should consider the most general $\binom{2}{0}$-tensor density, $\mathfrak{g}^{\mu\nu}$,  built with the available fields, and use it to build Lagrangian densities for the matter content. Following our precept, the matter content ---scalar matter--- should couple to $\mathfrak{g}^{\mu\nu}$, given by
\begin{dmath}
  \mathfrak{g}^{\mu\nu} = \alpha \, \nabla_\lambda T^{\mu,\nu\lambda} + \beta \, A_\lambda T^{\mu,\nu\lambda} + \gamma \, \epsilon_{\lambda\kappa\rho\sigma} T^{\mu, \lambda\kappa} T^{\nu, \rho\sigma},
  \label{geng}
\end{dmath}
with $\alpha$, $\beta$ and $\gamma$ arbitrary coefficients.

@@

It is well-known that this system of equations can be obtained effectively from the Einstein--Hilbert action coupled minimally to a massless scalar field
\begin{equation}
  S_{\text{eff}} = \frac{1}{\alpha} \int \dn{4}{x} \sqrt{g} \left( R + 2 \Lambda - \frac{\alpha}{2} g^{\mu\nu} \pa{\mu}\phi \pa{\nu} \phi \right).
\end{equation}
We have show, so far, that our model of polynomial purely affine gravity is compatible with General Relativity in the sector of vanishing torsion, and in addition the compatibility can be extended to include the interaction with a free scalar field. The analysis of this coupled system could be enhanced to include a potential for the scalar field,

@@

Longitudinal and transverse projectors can be introduced in order to decompose the action and specify in a more standard fashion the dynamics of Eq.~\eqref{4dmax}. This can be done since we have the metric $g_{\mu\nu}$ compatible with the connection, and $\nabla^\mu=g^{\mu\nu}\nabla_\nu$. Additionally we will denote by $\Delta = \nabla^\mu\nabla_\mu$ the covariant Laplacian. Let us define $P_{\texttt{L}}{}^\mu{}_\nu=\nabla^\mu\Delta^{-1} \nabla_\nu $, the projector into logitudinal vectors, and similarly $P_\perp{}^\mu{}_\nu=\delta^\mu_\nu -P_{\texttt{L}}{}^\mu{}_\nu$ is the projector into transverse vectors. Therefore, transverse indices are defined such that $\nabla_\mu A_\perp{}^{\mu}=0$.

Thus, the perturbations can be split into their longitudinal and transverse components as
\begin{dmath}
  a_\mu = {a_\perp}_{\mu} + {\nabla}_{\mu}{aL},
\end{dmath}
\begin{dmath}
  {t}^{\lambda \mu \nu} =
  {t_\perp}^{\lambda \mu \nu}
  + 2  {\nabla}^{\lambda} {t_{\texttt{a}}^{\mu \nu}}
  - 2  {\nabla}^{[\mu}{{t_{\texttt{a}}}^{\nu] \lambda}} 
  + 2  {\nabla}^{\mu}{{t_{\texttt{s}}}^{\nu \lambda}}
  + 2  {\nabla}^{\lambda [\mu}{{t_{\texttt{v}}}^{\nu]}}
\end{dmath}
and
\begin{dmath}
  {\gamma}^{\lambda}\,_{\mu \nu}  =
  {\gamma_\perp}^{\lambda}{}_{\mu \nu}
  + {\nabla}^{\lambda}{{\gamma_{\texttt{s}}}_{\mu \nu}}
  + 2  {\nabla}_{(\mu}{{\gamma}^{\lambda}\,_{\nu)}}
  + 2  {\nabla}^{\lambda}\nabla_{(\mu}{{\gamma_{\texttt{v}}}_{\nu)}}
  + 2  \nab{(\mu} \nab{\nu)} {{\gamma_{\texttt{v}'}}^{\lambda}} 
  + 2  {\nabla}^{\lambda}\nab{(\mu} \nab{\nu)}{\gamma_{\texttt{L}}},
\end{dmath}
where all the indices are transverse except those of the derivative $\nabla_\mu$, the subindex $\perp$ indicates that it corresponds to the fully transverse projection of the tensor in all of its indices, the subindex ${\texttt{L}}$ indicates that this is the longitudinal projection of the tensor, while the sub indices $\texttt{s}$, $\texttt{a}$ and $\texttt{v}$ specify that the tensor is symmetric, antisymmetric or a vector representation. In particular, $t_\perp^{\lambda}{}_\kappa$ is traceless. \hl{check the last sentence. Is this ${t_s}^{\alpha}{}_{\alpha}$?... If true, this condition will kill the scalars contribution in the next equations}
  \label{pert-Lambda}
    S = \int \dn{4}{x} \Bigg(
    C_1 \Big( {t_s}^{\alpha\beta }-\frac{1}{4}g^{\alpha\beta}{{t_s}^{\gamma}{}_{\gamma}} \Big) \bigg( -  (\frac{4\Lambda^2}{3}+\Lambda \Delta+\Delta^2) {\gamma_s}_{\alpha\beta}  
    - \frac{8}{3}   \Lambda\Delta {{\gamma}_{\alpha\beta}} 
    \bigg)
    + {{t_v}^{\alpha}}\bigg(
    \Big(( C_1 - 2\, C_2) \Delta^3 + \frac{1}{3}( C_1 + 2C_2 )\Lambda\Delta^2 + \frac{2}{3}( 2C_1 +  C_2 )\Lambda^2\Delta - \frac{1}{3}C_1  \Lambda^3\Big){{\gamma_v}_{\alpha}}
    + \Big((C_1 - 2\, C_2) \Lambda\Delta - 2\, C_2 \Delta^2\Big){{\gamma_\perp}^{\beta}\,_{\beta \alpha}}
    + 8\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} \Delta{\nabla}^{\beta }\Delta{{t_a}^{\gamma\delta}}  
    +\Big(- 2\, E_2 \Delta^2 +  (- E_2 + B_3 - \frac{2}{3}\, B_4) \Lambda\Delta\Big){a_\perp}_{\alpha} 
    \bigg)
    + {{t_s}^{\alpha}{}_{\alpha}}\bigg( \frac{2}{3}C_1 \Lambda\Delta{{\gamma}^{\beta}\,_{\beta}} 
    + \frac{2}{3}E_2 \Lambda \Delta{a_L}  
    - \frac{4}{3}C_1 \Lambda^2 \Delta{g_L}
    - \frac{1}{4}C_1 (\Lambda + \Delta)  \Delta{{\gamma_s}^{\beta}{}_{\beta}}
    \bigg)   \Bigg),
  \hl{1.- Is it Ok? 2.- Are the $B_3$ and $B_4$ there?}
  where $\Delta \equiv \nabla^{\alpha}\nabla_{\alpha}$ is the Laplacian operator.
  This action sets the dynamics of two transverse-traceless tensors (or specific combination of tensors), as well as two transverse vectors and two scalars.
%% We define the projector into logitudinal vectors $P_{\texttt{L}}{}^\mu_\nu = \tfrac{1}{p^2}p^\mu p_\nu $, with $p^\mu = g^{\mu\nu} p_\nu$, such that $$A_{\texttt{L}\mu} = P_{\texttt{L}}{}^\nu_\mu \, A_\nu = A_{\texttt{L}} \, p_{\mu},$$ where $A_{\texttt{L}}=\tfrac{1}{p^2} p^\mu A_\mu$. Similarly, $P_\perp{}^\mu_\nu=\delta^\mu_\nu -P_{\texttt{L}}{}^\mu_\nu$ defines the projector into transverse vectors, i.e., $A_{\perp \mu}=P_{\perp}{}_\mu^\nu A_\nu$. Thus,
%% \begin{dmath}
%%   \label{transversea}
%%   \tilde a_\nu=a_\perp{}_{\nu}+a_{\texttt{L}} p_\nu,
%% \begin{dmath}
%%   \label{transverset}
%%   \tilde t^{\lambda\mu\nu}=\frac{1}{p}\epsilon^{\mu\nu\kappa\rho}p_\rho t_\perp^{\lambda}{}_\kappa+\frac{1}{p^2}p^\lambda\epsilon^{\mu\nu\kappa\rho}p_\rho t_{\texttt{a}\kappa}-\frac{2}{p^2}p^{[\mu}\epsilon^{\nu]\lambda\kappa\rho}p_\rho t_{\texttt{a}\kappa}+\frac{2}{p}p^{[\mu} t_{\texttt{s}}^{\nu]\lambda}+\frac{2}{p^2}p^\lambda p^{[\mu} t_{\texttt{v}}^{\nu]},
%% \begin{dmath}
%%   \label{transversegamma}
%%   \tilde \gamma^{\lambda}{}_{\mu\nu}=\gamma_\perp^{\lambda}{}_{\mu\nu}+\frac{1}{p}p^\lambda \gamma_{\texttt{s}}{}_{\mu\nu}+\frac{2}{p}p_{(\mu} \gamma^\lambda{}_{\nu)}+\frac{2}{p^2}p^\lambda p_{(\mu} \gamma_{\texttt{v}}{}_{\nu)}+ \frac{1}{p^2}p_{\mu} p_{\nu} \gamma'_{\texttt{v}}{}^\lambda+\frac{1}{p^3}p^\lambda p_{\mu} p_{\nu} \gamma_{\texttt{L}},
%% where all the indices are transverse except those of the momentum $p_\mu$, the subindex $\perp$ indicates that it corresponds to the fully transverse projection of the tensor in all of its indices, the subindex ${\texttt{L}}$ indicates that this is the longitudinal projection of the tensor, while the subindices $\texttt{s}$, $\texttt{a}$ and $\texttt{v}$ specify that the tensor is symmetric, antisymmetric or a vector representation. In particular, $t_\perp^{\lambda}{}_\kappa$ is traceless. Additionally, we could separate the fields into irreducible representations by projecting them into their traceless part and other lower dimensional tensor representations. In the meanwhile we shall not do this expansion.
%% \begin{widetext}
%%   Using Eqs.~\eqref{transversea}, \eqref{transverset} and~\eqref{transversegamma}, the action~\eqref{4dflat} is then rewritten
%%   \begin{dmath}
%%     \label{4dtransverse}
%%     \int\dn{4}{p}p^2
%%     \bigg[
%%       2C_1 \hat{p}_\mu \hat{p}_\sigma\Big(\gamma_\perp^{\mu}{}_{\rho\nu}+\hat{p}^\mu \gamma_{\texttt{s}}{}_{\rho\nu}+2\hat{p}_{(\rho} \gamma^\mu{}_{\nu)}+2\hat{p}^\mu \hat{p}_{(\rho} \gamma_{\texttt{v}}{}_{\nu)}+ \hat{p}_{\rho} \hat{p}_{\nu} \gamma'_{\texttt{v}}{}^\mu+\hat{p}^\mu \hat{p}_{\rho} \hat{p}_{\nu} \gamma_{\texttt{L}}\Big)^* \Big(\epsilon^{\rho\sigma\delta\eta}\hat{p}_\eta t_\perp^{\nu}{}_\delta+\hat{p}^\nu t_{\texttt{a}}^{\rho\sigma}-2\hat{p}^{[\rho} t_{\texttt{a}}^{\sigma]\nu}+2\hat{p}^{[\rho} t_{\texttt{s}}^{\sigma]\nu}+2\hat{p}^\nu \hat{p}^{[\rho} t_{\texttt{v}}^{\sigma]}\Big)
%%       +2C_2 \hat{p}_{\mu} \hat{p}_\sigma\Big(\gamma_\perp^{\rho}{}_{\nu\rho}+\hat{p}^\rho \gamma_{\texttt{s}}{}_{\nu\rho}+2\hat{p}_{(\nu} \gamma^\rho{}_{\rho)}+2\hat{p}^\rho \hat{p}_{(\nu} \gamma_{\texttt{v}}{}_{\rho)}+\hat{p}_{\nu} \hat{p}_{\rho} \gamma'_{\texttt{v}}{}^\rho +\hat{p}^\rho \hat{p}_{\nu} \hat{p}_{\rho} \gamma_{\texttt{L}}\Big)^* \Big(\epsilon^{\mu\nu\kappa\eta}\hat{p}_\eta t_\perp^{\sigma}{}_\kappa+\hat{p}^\sigma t_{\texttt{a}}^{\mu\nu}-2\hat{p}^{[\mu} t_{\texttt{a}}^{\nu]\sigma}+2\hat{p}^{[\mu} t_{\texttt{s}}^{\nu]\sigma}+2\hat{p}^\sigma \hat{p}^{[\mu} t_{\texttt{v}}^{\nu]}\Big)
%%       +E_1 \hat{p}_{\rho} \hat{p}_{\sigma}\Big(\epsilon^{\mu\nu\kappa\eta}\hat{p}_\eta t_\perp^{\rho}{}_\kappa+\hat{p}^\rho t_{\texttt{a}}^{\mu\nu}-2\hat{p}^{\mu} t_{\texttt{a}}^{\nu\rho}+2\hat{p}^{\mu} t_{\texttt{s}}^{\nu\rho}+2\hat{p}^\rho \hat{p}^{\mu} t_{\texttt{v}}^{\nu}\Big)^*\Big(\epsilon^{\lambda\kappa\delta\eta}\hat{p}_\eta t_\perp^{\sigma}{}_\delta+\hat{p}^\sigma t_{\texttt{a}}^{\lambda\kappa}-2\hat{p}^{\lambda} t_{\texttt{a}}^{\kappa\sigma}+2\hat{p}^{\lambda} t_{\texttt{s}}^{\kappa\sigma}+2\hat{p}^\sigma \hat{p}^{\lambda} t_{\texttt{v}}^{\kappa}\Big)\epsilon_{\mu\nu\lambda\kappa}
%%       +E_2 \hat{p}_{\lambda} \hat{p}_{\mu} \Big(\epsilon^{\mu\nu\kappa\rho}\hat{p}_\rho t_\perp^{\lambda}{}_\kappa+\hat{p}^\lambda t_{\texttt{a}}^{\mu\nu}-2\hat{p}^{[\mu} t_{\texttt{a}}^{\nu]\lambda}+2\hat{p}^{[\mu} t_{\texttt{s}}^{\nu]\lambda}+2\hat{p}^\lambda \hat{p}^{[\mu} t_{\texttt{v}}^{\nu]}\Big)^*\Big(a_\perp{}_\nu+\hat{p}_\nu a_{\texttt{L}}\Big)
%%       \bigg],
%% \end{widetext}
%% simplifiying
%% \begin{dmath}
%%   \label{WithGhosts}
%%   S=\int d^4p\Bigg[-2C_1 p^2\gamma_{\texttt{s}\mu\nu}^*t_{\texttt{s}}^{\mu\nu}
%%     +p^2\Big((-2C_1+2C_2)\gamma_{\texttt{v}\mu}+2C_2\gamma_{\perp\mu\nu}^\nu
%%     +8E_1t_{\texttt{a}\mu}+E_2a_{\perp\mu}\Big)^*t_{\texttt{v}}^\mu
%%     \Bigg]
When we restrict ourselves to perturbations around a flat (Minkowski) spacetime, the above equation simplifies greatly,
  S = \int \dn{4}{x} \, \bigg[
    - C_1 ( \Delta\gamma_{\texttt{s}\mu\nu})^* \Delta t_{\texttt{s}}^{\mu\nu}
    + \Big( ( C_1 - 2 C_2) \nabla^\nu \Delta \nabla_\nu \gamma_{\texttt{v}\mu}
    - 2 C_2 \Delta \gamma^\nu_{\perp\nu\mu}
    - 8 E_1 \epsilon_{\alpha\beta\nu\mu} \nabla^{\alpha} \Delta t_{\texttt{a}}^{\beta\nu}
    - 2 E_2 \Delta a_{\perp\mu} \Big)^* \Delta t_{\texttt{v}}^\mu
    \bigg],
\hl{Where are the scalars? This has 18 dof, while the previous one has 20 dof. According to the counting, there are 18... I suppose that eighteen is the correct answer, but What is wrong in the first expression?} and it is clear that there are only eighteen degrees of freedom, corresponding to a couple of transverse-traceless rank-two-tensors, and a couple of transverse vectors.

@@

>>>>

%% --------- Abstract
\begin{abstract}
  We show that the effective field equations for a recently formulated polynomial affine model of gravity, in the sector of a torsion-free connection, accept general Einstein manifolds---with or without cosmological constant---as solutions. Moreover, the effective field equations coincide with some of those obtained from a gravitational Yang--Mills theory known as Stephenson--Kilmister--Yang theory. Additionally, we find a generalization of a minimally coupled massless scalar field in General Relativity with a ``minimally'' coupled scalar field in this affine model. Finally, we present a naive analysis of the propagators of the gravitational theory, an their relation with the presence of ghost in the model. For completeness we include three appendices, including one in which it is shown that a Birkhoff-like theorem is valid for the analyzed sector.
\end{abstract}

@@

However, General Relativity seems to be an effective theory of a, yet to be discovered, more fundamental one. A first hint was pointed out by E.~Cartan, after remarking that the curvature is not built up from a metric field but from a connection, and that a part of it was left behind if one chooses to work with the Levi-Civita connection~\cite{Cartan1922,Cartan1923,Cartan1924,Cartan1925}. One could also see this by considering Palatini's approach in General Relativity, in which both metric and connection fields are independent, and noticing that generically the theory possesses terms which do not appear in General Relativity. In addition, we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization, General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}, suggests the theory can be thought as an effective theory. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, by Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

%% Classically, General Relativity is the most successful theory of gravitational interactions. However, it is believed to be the effective theory of a, yet to be discovered, more fundamental one. Among the problems of General Relativity we mention a few aspects related with the quantization: First, the fact that in the standard formalism of quantization General Relativity is non-renormalizable~\cite{'tHooft:1973us,'tHooft:1974bx,Deser:1974cz,Deser:1974cy}. Second, the Wheeler--DeWitt equations are not well-defined in general due to their nonpolynomial dependence on the Arnowitt--Deser--Misner (or ADM) variables~\cite{Arnowitt:1959ah,Arnowitt:1960es,WheelerGeo,DeWitt:1967yk,DeWitt:1967ub,DeWitt:1967uc}. One of the most successful quantization methods of gravitational interactions---developed from ideas by Gambini and Trias~\cite{Gambini:1980yz,Gambini:1986ew}, and Ashtekar~\cite{Ashtekar:1986yd,Ashtekar:1987gu}---, hints that General Relativity is not as complete as expected, since the Lagrangian formulation includes (at least) an extra term known as Holst term~\cite{Holst:1995pc}. On the other hand, it is well known that in the standard formulation of General Relativity the skew-symmetric part of the connection (or torsion) is assumed to be zero, albeit it is known that such assumption can be relaxed once matter is added, in particular fermionic matter~\cite{Kibble:1961ba,Hehl:1976kj,Shapiro:2001rz,Hammond:2002rm}.

@@

Equation~\eqref{SimpleEOM} is a generalization of a condition known in Riemannian geometry as covariantly constant Ricci curvature---aka parallel Ricci curvature---, \mbox{$\nab{\rho} R_{\mu\nu} = 0$.}  All Einstein manifolds, whose Ricci tensor is proportional to the metric, \mbox{$R_{\mu\nu} \propto g_{\mu\nu}$,} satisfy the parallel Ricci condition due to the metricity condition, and therefore every vacuum solution to the Einstein's equations solves the (torsion-free) field equations of our model. Consequently, the fact that the non-relativistic limit of the gravitational potential in Eq.~\eqref{NewtonPot} yields a Newtonian potential seems clearer, and we can argue that even the post-Newtonian corrections coming from General Relativity are present in the chosen scenario of our model. An additional comment, the parallel Ricci curvature assures that although the manifold is not Einstein in general, the metric has to be locally a product of Einstein metrics~\cite{Besse}.

@@

Interestingly, the Eq.~\eqref{SimpleEOM}---equivalently Eq.~\eqref{harm-curv}---can be obtained from an effective gravitational Yang--Mills functional action~\cite{stephenson1958quadratic,kilmister1961use,Yang1974}, %constructed as the trace of the 
\begin{equation}
  \label{SKY}
  S_{\textsc{ym}} = \int \Tr \left( \Rif{}{} \st \Rif{}{} \right) = \int \left( \Rif{a}{b} \st \Rif{b}{a} \right),
\end{equation}
where $\Rif{}{} \in \Omega^2(\Mi, T^*\Mi \otimes T\Mi)$ is the curvature two-form, the operator $\st$ denotes the Hodge star, and the trace is taken on the bundle indices (see Ref.~\cite{bourguignon1982yang}).

@@

Until now, the most general diffeomorphism-invariant and power-counting renormalizable (gravitational) theory for an affine connection has been built, and we have showed that in a certain sector it is equivalent to General Relativity. In what follows, we show an attempt of including scalar matter into the model. The theory does not require a spacetime metric, but instead we should consider the most general $\binom{2}{0}$-tensor density, $\mathfrak{g}^{\mu\nu}$,  built with the available fields, and use it to build Lagrangian densities for the matter content. Following our precept, the matter content---scalar matter---should couple to $\mathfrak{g}^{\mu\nu}$, given by
\begin{dmath}
  \mathfrak{g}^{\mu\nu} = \alpha \, \nabla_\lambda T^{\mu,\nu\lambda} + \beta \, A_\lambda T^{\mu,\nu\lambda} + \gamma \, \epsilon_{\lambda\kappa\rho\sigma} T^{\mu, \lambda\kappa} T^{\nu, \rho\sigma},
  \label{geng}
\end{dmath}
with $\alpha$, $\beta$ and $\gamma$ arbitrary coefficients.

@@

It is well-known that this system of equations can be obtained effectively from the Einstein--Hilbert action coupled minimally to a massless scalar field
\begin{equation}
  S_{\text{eff}} = \frac{1}{\alpha} \int \dn{4}{x} \sqrt{g} \left( R + 2 \Lambda - \frac{\alpha}{2} g^{\mu\nu} \pa{\mu}\phi \pa{\nu} \phi \right).
\end{equation}
We have show, so far, that our model of polynomial purely affine gravity is compatible with General Relativity in the sector of vanishing torsion, and in addition the compatibility can be extended to include the interaction with a free scalar field. The analysis of this coupled system could be enhanced to include a potential for the scalar field, however, we are not considering this case at the moment.

@@

Longitudinal and transverse projectors can be introduced in order to decompose the action and specify in a more standard fashion the dynamics of Eq.~\eqref{4dmax}. This can be done since we have the metric $g_{\mu\nu}$ compatible with the connection, and $\nabla^\mu=g^{\mu\nu}\nabla_\nu$. Additionally we will denote by $\Delta = \nabla^\mu\nabla_\mu$ the covariant Laplacian. Let us define $P_{\mathtt{L}}{}^\mu{}_\nu=\nabla^\mu\Delta^{-1} \nabla_\nu $, the projector into logitudinal vectors, and similarly $P_\perp{}^\mu{}_\nu=\delta^\mu_\nu -P_{\mathtt{L}}{}^\mu{}_\nu$ is the projector into transverse vectors. Therefore, transverse indices are defined such that $\nabla_\mu A_\perp{}^{\mu}=0$.

Thus, the perturbations can be split into their longitudinal and transverse components as
\begin{dmath}
  a_\mu = {a_\perp}_{\mu} + {\nabla}_{\mu}{a_{\mathtt{L}}},
\end{dmath}
\begin{dmath}
  {t}^{\lambda \mu \nu} =
  {t_\perp}^{\lambda \mu \nu}
  + 2  {\nabla}^{\lambda} {t_{\mathtt{a}}^{\mu \nu}}
  - 2  {\nabla}^{[\mu}{{t_{\mathtt{a}}}^{\nu] \lambda}} 
  + 2  {\nabla}^{\mu}{{t_{\mathtt{s}}}^{\nu \lambda}}
  + 2  {\nabla}^{\lambda [\mu}{{t_{\mathtt{v}}}^{\nu]}}
\end{dmath}
and
\begin{dmath}
  {\gamma}^{\lambda}\,_{\mu \nu}  =
  {\gamma_\perp}^{\lambda}{}_{\mu \nu}
  + {\nabla}^{\lambda}{{\gamma_{\mathtt{s}}}_{\mu \nu}}
  + 2  {\nabla}_{(\mu}{{\gamma}^{\lambda}\,_{\nu)}}
  + 2  {\nabla}^{\lambda}\nabla_{(\mu}{{\gamma_{\mathtt{v}}}_{\nu)}}
  + 2  \nab{(\mu} \nab{\nu)} {{\gamma_{\mathtt{v}'}}^{\lambda}} 
  + 2  {\nabla}^{\lambda}\nab{(\mu} \nab{\nu)}{\gamma_{\mathtt{L}}},
\end{dmath}
where all the indices are transverse except those of the derivative $\nabla_\mu$, the subindex $\perp$ indicates that it corresponds to the fully transverse projection of the tensor in all of its indices, the subindex ${\mathtt{L}}$ indicates that this is the longitudinal projection of the tensor, while the sub indices $\mathtt{s}$, $\mathtt{a}$ and $\mathtt{v}$ specify that the tensor is symmetric, antisymmetric or a vector representation.
%%In particular, $t_\perp^{\lambda}{}_\kappa$ is traceless. \hl{check the last sentence. Is this ${t_s}^{\alpha}{}_{\alpha}$?... If true, this condition will kill the scalars contribution in the next equations}

\begin{widetext}
  The perturbed action is then
  \begin{dmath}[compact, spread=2pt]
  \label{pert-Lambda}
    S = \int \dn{4}{x} \Bigg(
    C_1 \Big( {t_{\mathtt{s}}}^{\alpha\beta }-\frac{1}{4}g^{\alpha\beta}{{t_{\mathtt{s}}}^{\gamma}{}_{\gamma}} \Big) \bigg( -  (\frac{4\Lambda^2}{3}+\Lambda \Delta+\Delta^2) {\gamma_{\mathtt{s}}}_{\alpha\beta}  
    - \frac{8}{3}   \Lambda\Delta {{\gamma}_{\alpha\beta}} 
    \bigg)
    + {{t_{\mathtt{v}}}^{\alpha}}\bigg(
    \Big(( C_1 - 2\, C_2) \Delta^3 + \frac{1}{3}( C_1 + 2C_2 )\Lambda\Delta^2 + \frac{2}{3}( 2C_1 +  C_2 )\Lambda^2\Delta - \frac{1}{3}C_1  \Lambda^3\Big){{\gamma_{\mathtt{v}}}_{\alpha}}
    + \Big((C_1 - 2\, C_2) \Lambda\Delta - 2\, C_2 \Delta^2\Big){{\gamma_\perp}^{\beta}\,_{\beta \alpha}}
    + 8\, E_1 {\epsilon}_{\alpha \beta \gamma \delta} \Delta{\nabla}^{\beta }\Delta{{t_{\mathtt{a}}}^{\gamma\delta}}  
    +\Big(- 2\, E_2 \Delta^2 +  (- E_2 + B_3 - \frac{2}{3}\, B_4) \Lambda\Delta\Big){a_\perp}_{\alpha} 
    \bigg)
    + {{t_{\mathtt{s}}}^{\alpha}{}_{\alpha}}\bigg( \frac{2}{3}C_1 \Lambda\Delta{{\gamma}^{\beta}\,_{\beta}} 
    + \frac{2}{3}E_2 \Lambda \Delta{a_{\mathtt{L}}}  
    - \frac{4}{3}C_1 \Lambda^2 \Delta{g_{\mathtt{L}}}
    - \frac{1}{4}C_1 (\Lambda + \Delta)  \Delta{{\gamma_{\mathtt{s}}}^{\beta}{}_{\beta}}
    \bigg)   \Bigg),
  \end{dmath}
  \hl{1.- Is it Ok?}
\end{widetext}
where $\Delta \equiv \nabla^{\alpha}\nabla_{\alpha}$ is the Laplacian operator. This action sets the dynamics of two transverse-traceless tensors (or specific combination of tensors), as well as two transverse vectors and two scalars, which are associated to eighteen degrees of freedom.

@@

In addition, when we restrict ourselves to perturbations around a flat (Minkowski) spacetime, the above equation simplifies greatly,
\begin{dmath}
  S = \int \dn{4}{x} \, \bigg[
    - C_1 ( \Delta\gamma_{\mathtt{s}\mu\nu})^* \Delta t_{\mathtt{s}}^{\mu\nu}
    + \Big( ( C_1 - 2 C_2) \nabla^\nu \Delta \nabla_\nu \gamma_{\mathtt{v}\mu}
    - 2 C_2 \Delta \gamma^\nu_{\perp\nu\mu}
    - 8 E_1 \epsilon_{\alpha\beta\nu\mu} \nabla^{\alpha} \Delta t_{\mathtt{a}}^{\beta\nu}
    - 2 E_2 \Delta a_{\perp\mu} \Big)^* \Delta t_{\mathtt{v}}^\mu
    \bigg],
\end{dmath}
and it is clearer that there are only eighteen degrees of freedom.
